# Introduction

## What you will learn

## How book is organised 

By data
Code interwoven

## Prerequisites (Content drawn from existing resources such as https://r4ds.hadley.nz/intro#prerequisites)  

### R 
Download 
Point to intro to R content (RUWithme, Environmental Computing, Software Carpentry)
	
### RStudio 
RStudio projects
Point to resource about Rproj (SWC)

Running R code
https://r4ds.hadley.nz/intro#running-r-code

### Version control with git

### R packages

Every code section will always begin with calls to R packages. There will be commented out code for you to install these. 

Here are few packages we consider essential to working with Psychology data. 

- tidyverse 
- ggplot2
- here
- janitor

At the end of each chapter, we will also include our call to `sessionInfo()` so you can see what version of which packages we are using. 

### Virtual environments

Speaking on what package versions we are using as we write this book, we understand the R package space is constantly changing. This means sometimes code will break due to package updates and this is all part of the process! To combat this problem, we've enlisted [`renv`](https://rstudio.github.io/renv/articles/renv.html#getting-started) to create a reproducible environment for building this book. 

#### Download our virtual environment

The virtual environment used to build this book is stored in a `lockfile`

#### Install

```{r}
install.packages("renv")

library(renv)
```
#### Restore

```{r, eval=FALSE}
restore()
```





